### 2012-12-16 course11/descr/src/latex/CourseReport-Adaptor.tex
Changed tree layout. Added report to tree.

\section{Инструментарий коллективной оптимизации программ}
% TODO