### 2013-01-03 course11/descr/src/latex/CourseReport-Task.tex
Исправлены мелочи в РПЗ

\section*{Учебное задание}
% TODO
Задание

\pagebreak