### 2013-02-05 practice12/descr/src/latex/PracticeReport-Adaptor.tex
Продолжена переделка старого документа.

\section{Конструкторская часть}