### 2013-02-05 practice12/descr/src/latex/PracticeReport-Experiment.tex

\section{Экспериментальная часть}