### 2013-02-05 practice12/descr/src/latex/PracticeReport-Research.tex

\section{Исследовательская часть}