### 2013-03-12 practice12/descr/src/latex/PracticeReport-Adaptor.tex
Established basic structure.

\part{Конструкторская часть}

\section{Реализация модели производительности программ}