### 2013-03-12 practice12/descr/src/latex/PracticeReport-Experiment.tex

\part{Экспериментальная часть}