### 2013-03-12 practice12/descr/src/latex/PracticeReport-Research.tex

\part{Исследовательская часть}
\section{Моделирование производительности программ}
\section{Библиотеки машинного обучения для языка Python}
\subsection{Рассмотрение альтернатив}
\subsubsection{Orange}
\subsubsection{sklearn}
\subsubsection{PyML}
\subsubsection{mlpy}