### 2013-03-12 practice12/descr/src/latex/PracticeReport-Theory.tex

\part{Теоретическая часть}

\section{Регрессионный анализ}

Hello \cite{lag}.