### 2013-05-07 diploma/src/latex/DiplomaProject-Appendix.tex
Added a named report and extended evaluation part.

\section*{Приложение}
\addcontentsline{toc}{section}{Приложение}%

\begin{figure}[H]
    \center{\includegraphics[width=0.6\linewidth]{2-Attribute-Statistics-time}}
    \caption{Графическое изображение распределения свойства \texttt{time}.}
    \label{img:2-Attribute-Statistics-time}
\end{figure}