### 2013-09-23 doc/latex/Article.tex

\documentclass{disser}

\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{adjustbox}

\graphicspath{{../pictures/}}

\begin{document}

\input{Article-Abstract}

\input{Article-Introduction}

\input{Article-Methodology}

\input{Article-Construction}

\input{Article-Evaluation}

\input{Article-Conclusion}

\end{document}