### 2013-09-30 keynote/Keynote.tex
Translate existing keynote to English

\documentclass{beamer}
\usepackage[T2A]{fontenc}
\usepackage[english,russian]{babel}
\usepackage[utf8x]{inputenc}

\usetheme{Copenhagen}

\title{Programs Performance Analysis Toolkit}
\author{Michael Pankov}
\institute{Bauman Moscow State Technical University}
\date{June 19, 2013}

\makeatletter
%\definecolor{beamer@blendedblue}{rgb}{0.5,0.5,0.3} % changed this
\definecolor{beamer@darkgreen}{RGB}{35,150,35}

%\setbeamercolor{normal text}{fg=black,bg=white}
%\setbeamercolor{alerted text}{fg=red}
%\setbeamercolor{example text}{fg=green!50!black}

\setbeamercolor{structure}{fg=beamer@darkgreen}

%\setbeamercolor{background canvas}{parent=normal text}
%\setbeamercolor{background}{parent=background canvas}

\setbeamercolor{palette primary}{bg=beamer@darkgreen} % changed this
\setbeamercolor{palette secondary}{use=structure,fg=structure.fg!100!green} % changed this
\setbeamercolor{palette tertiary}{use=structure,fg=structure.fg!100!green} % changed this
\makeatother

\begin{document}

\maketitle

\begin{frame}
\frametitle{Goal, tasks, and importance of the work}

\begin{block}{Goal}
	\begin{itemize}
		\item Develop a method and software toolkit for modeling of programs performance on general purpose computers
	\end{itemize}
\end{block}

\begin{block}{Tasks}
	\begin{enumerate}
		\item Develop method of programs performance modeling 
		\item Implement the programs performance analysis \& modeling toolkit
		\item Study the efficiency of programs performance modeling toolkit on a set of benchmarks
	\end{enumerate}
\end{block}

\begin{block}{Importance}
	\begin{enumerate}
		\item Estimation of computer performance during its design (see co-design)
		\item Search of optimal compiler settings by methods of iterative compilation and machine learning-driven techniques
	\end{enumerate}
\end{block}

\end{frame}

\end{document}