### 2013-09-30 keynote/Keynote.tex
Add 2 slides on method

\documentclass{beamer}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}

\usetheme{Copenhagen}

\title{Programs Performance Analysis Toolkit}
\author{Michael Pankov}
\institute{Bauman Moscow State Technical University}
\date{June 19, 2013}

\makeatletter
%\definecolor{beamer@blendedblue}{rgb}{0.5,0.5,0.3} % changed this
\definecolor{beamer@darkgreen}{RGB}{35,150,35}

%\setbeamercolor{normal text}{fg=black,bg=white}
%\setbeamercolor{alerted text}{fg=red}
%\setbeamercolor{example text}{fg=green!50!black}

\setbeamercolor{structure}{fg=beamer@darkgreen}

%\setbeamercolor{background canvas}{parent=normal text}
%\setbeamercolor{background}{parent=background canvas}

\setbeamercolor{palette primary}{bg=beamer@darkgreen} % changed this
\setbeamercolor{palette secondary}{use=structure,fg=structure.fg!100!green} % changed this
\setbeamercolor{palette tertiary}{use=structure,fg=structure.fg!100!green} % changed this
\makeatother

% Multi-slide enumeration
% http://tex.stackexchange.com/questions/55000/continuing-enumerate-counters-in-beamer
\newcounter{saveenumi}
\newcommand{\seti}{\setcounter{saveenumi}{\value{enumi}}}
\newcommand{\conti}{\setcounter{enumi}{\value{saveenumi}}}

\resetcounteronoverlays{saveenumi}

\begin{document}

\maketitle

\begin{frame}
\frametitle{Goal, tasks, and importance of the work}

\begin{block}{Goal}
	\begin{itemize}
		\item Develop a method and software toolkit for modeling of programs performance on general purpose computers
	\end{itemize}
\end{block}

\begin{block}{Tasks}
	\begin{enumerate}
		\item Develop method of programs performance modeling 
		\item Implement the programs performance analysis \& modeling toolkit
		\item Study the efficiency of programs performance modeling toolkit on a set of benchmarks
	\end{enumerate}
\end{block}

\begin{block}{Importance}
	\begin{enumerate}
		\item Estimation of computer performance during its design (see co-design)
		\item Search of optimal compiler settings by methods of iterative compilation and machine learning-driven techniques
	\end{enumerate}
\end{block}

\end{frame}

\begin{frame}
\frametitle{Overview}

	\begin{itemize}
		\item A lot of recent research: see C.\,Dubach, G.\,Fursin, B.\,C.\,Lee, W.\,Wu
		\item In particular, there's \textit{cTuning} public repository for research and corresponding program \textit{Collective Mind} run by G.\,Fursin
		\item This work is about modeling of performance of general-purpose computer programs with feature ranking by means of \textit{Earth Importance} and regression by means of \textit{k-Nearest Neighbors} and \textit{Earth Regression}. We try to accomplish automatic detection of relevant features.
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Method of statistical programs performance analysis ''Velocitas''}

	\begin{enumerate}
		\item Perform a series of experiments on measuring time of program execution and form a set, $U$:
			\begin{equation*}
				U = \{ ( X_i, y_i ) \}, X_i = (x_{ij}, i \in [1;m], j \in [1;n])
			\end{equation*}
		$X_i$ --- features vector (CPU frequency, number of rows of processed matrix, etc.), $y_i$ --- response (execution time), $m$ --- number of experiments, $n$ --- number of features
		\seti
	\end{enumerate}

\end{frame}

\begin{frame}
	\begin{enumerate}
		\conti
		\item Split the $U$ set into training sample $D$ and test sample $C$ by randomly assigning of 70\% of experiments to $D$
			\begin{eqnarray*}
				D &=& \{d_i | f_{rand}^I(d_i) > 0.3\}, \\
				d_i &=& (X_i, y_i), \\
				f_{rand}^I(d) &\in & [0:1], \\
				i &\in & [1;m], \\
				C &=& U \setminus D
			\end{eqnarray*}
		\seti
	\end{enumerate}
\end{frame}

\end{document}